Learning a reward function from demonstrations: a cascaded supervised learning approach

### 2012-05-21 piot2012cascading/main_KLEIN_v3.tex
Amélioration de détails dans le papier et nouvelles courbes pour Matthieu

\title{A cascading approach}
\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
(if needed)\\

@@

Given a sequential decision making problem posed in the MDP formalism, it is often difficult to specify manually a reward function in order to realize a given task. However, observing an expert demonstrating a task is often possible and more easier~\cite{ng2000algorithms}. That is why learning from demonstrations and more particularly IRL is an interesting problem. The IRL framework, fisrt introduced in~[\cite{russell1998learning},\cite{ng2000algorithms}], assumes that an expert, demonstrating a task, is acting optimally in a MDP with respect to an unknown reward function to be discovered. Unlike direct methods of learning from demonstrations which learn a mapping from states to actions via a supervised learning~[\cite{atkeson1997robot},\cite{pomerleau1989alvinn}], IRL methods must recover a reward function which is the most compact and tranferable way to specify a task. Moreover most of the direct methods do not use the structure of the MDP (one notable excepetion is~\cite{melo2010learning}), hence mimic the policy of the expert and are not able to generalize correctly the expert policy outside state examples. In order to overcome this drawback, undirect methods known as apprenticeship learning were introduced: the first one is~\cite{ng2004feature}. These methods (see~\cite{neu2009training} for a good survey) uses the IRL framework but most of them assume the dynamic of the environment known, or an environment model (via a simulator) must be available so as to test the effects of a policy. Moreover, most of existing undirect approaches require to solve the direct reinforcement learning problem (find an optimal policy knowing a given reward) several times (one notable exception is~\cite{boularias2011relative}). Finally, in many cases, the output of the algorithm is a policy but not a reward function (see section \ref{section: related work} for an in-depth presentation). These constraints can lead to the development of algorithms unapplicable to real problems and make the problem of apprenticeship learning harder than reinforcement learning. That is why a reborn interest for direct methods which passively imitate the expert exists. More particularly the reduction of the apprenticeship learning to classification, first introduced in~\cite{zadrozny2003cost} and recently used in~\cite{melo2010learning}, is legitimated for finite horizons problems in~[\cite{syed2010reduction},\cite{ross2010efficient}]. Our approach tries to avoid the drawbacks of the direct and undirect methods by cascading two well known supervised learning(see section \ref{section: Cascading}): first a step of classification and then a step of regression which introduces the MDP structure and  outputs a reward function (see \ref{section: Cascading}). Our approach doesn't require any of the following: complete trajectories from an expert, a generative model of the environment, the knowledge of the transition probabilities, the ability to compute a (near)-optimal policy for different rewards and hence resolve the forward reinforcement learning (RL) problem, the expert's policy. In addition, only a small amount of expert demonstrations (not even in the form of trajectories but simple transitions) is required.\\
The remainder of the paper is oraganized as follows: section \ref{section: background} gives us the different notations used in the paper and presents briefly the RL and IRL frameworks, section \ref{section: Cascading} present our method which is legitimated in section \ref{section: Cascading}, \ref{section: experiments} presents an instanciation of our method and tests it on different benchmarks and finally section \ref{section: related work} compares our method to the state of the art.

@@

A (finite) MDP \cite{puterman1994markov} is a tuple $\{S,A,P,\gamma,R\}$ where $S=\{s_i\}_{1\leq i \leq N}$ is the finite state space with $N\in\mathbb{N}^*$ states, $A=\{a_k\}_{1\leq k \leq K}$ is the finite action space with $K\in\mathbb{N}^*$ actions, $P=\{P_{s,a}\}_{(s,a)\in S\times A}$ is the set of transition probabilities where $P_{s,a}$ is a distribution probability over $S$, $P$ represents the dynamics of the MDP and the notation $p(s,a,s')=P_{s,a}(s')$ is often used and quantifies the probability to reach $s'\in S$ knowing that the action $a \in A$ was taken in the state $s\in S$, $\gamma\in]0,1[$ is the discount factor and $R$ is a function from $S\times A$ to $\mathbb{R}$ called the reward function. A deterministic policy $\pi$ is a function from $S$ to $A$, $\Pi$ is the set of deterministic policies and the value function $V^\pi_R$ is a function from $S$ to $\mathbb{R}$ defined by:

@@

%Inversely, a real column-vector $X=[X_i]_{1\leq i \leq N}$ can be identified to the function from $S$ to $\mathbb{R}$ that maps $s_i$ to $X_i$ for all $i\in\{1,\dots,N\}$ and the real rectangular-matrice $Y=[Y_{i,k}]_{1\leq i \leq N, 1\leq k \leq K}$ can be identified to the function from $S\times A$ to $\mathbb{R}$ that maps $(s_i,a_k)$ to $Y_{i,k}$ for all $(i,k)\in\{1,\dots,N\}\times\{1,\dots,K\}$.
For a finite-MDP $\{S,A,P,\gamma,R\}$ the existence of an optimal policy for any function $R$ is shown in \cite{bertsekas2001dynamic}.

@@

We will also need the following matrices and column-vectors defined for any dynamic $P$, any policy $\pi$ and reward function $R$ :$P_a=[P_{s_i,a}(s_j)]_{1\leq i,j \leq N}$, $P_\pi=[P_{s_i,\pi(s_i)}(s_j)]_{1\leq i,j \leq N}$, $R_\pi=[R(s_i,\pi(s_i))]_{1\leq i,j \leq N}$. The transposition of a real column-vector $X=[X_i]_{1\leq i \leq M}$, $M\in\mathbb{N}^*$, is the real row-vector $X^T$ and $\|X\|_{\infty}=\max_{1\leq i \leq N}|X_i|$. The transposition of the real matrice $Y=[Y_{i,j}]_{1\leq i \leq I, 1\leq j \leq J}$, $(I,J)\in\mathbb{N}^2$, will be $Y^T$ and $\|Y\|_{\text{max}}=\max_{1\leq i \leq I,1\leq j \leq J}|Y_{i,j}|$.  Thanks to these vectorial notations, we can define the Bellman operator $T^\pi_R$, for a given deterministic policy $\pi$ and a given reward $R$, which is a function from $\mathbb{R}^N$ to $\mathbb{R}^N$ such that:

@@

representation), that the model (P and R) may be unknown (the only information being provided through rewarded transition sampled according
to some behavioral policy), that learning can occurs in an online or offline setting, and so on. There exists good books on the subject [\cite{bertsekas2001dynamic},\cite{sutton1998reinforcement}].\\
In the classical IRL paradigm \cite{ng2000algorithms}, an MDP without reward $\{S,A,P,\gamma\}$ and a policy $\pi_E$ called expert-policy are given and the problem is to find
a reward $R^*$ for which the policy $\pi_E$ will be optimal. However this problem is clearly ill-posed in the sense that there is no uniqueness of the reward $R^*$ : many rewards functions are equivalent in the sense that they have the same optimal deterministic policies \cite{ng1999policyreward}, moreover the trivial zero-reward is a solution for any deterministic policy $\pi_E$ as it is shown in \cite{ng2000algorithms}. In the literature, some solutions are proposed in order to respond to the ill-posed nature of the problem (\cite{ng2000algorithms},\cite{ziebart2008maximum},\cite{boularias2011relative}).
In this paper, we assume that the solely available information is provided by transitions sampled according to the expert policy $\pi_E$:

@@

\{(s_i,a_i=\pi_E(s_i)),s_i')\}_{1\leq i \leq D}, D\in\mathbb{N}^*.

@@

@@

Finally let's recall some basic but important results of finite-Markov chain theory.

@@

@@

The available data being depicted in \eqref{equation:data}, a first problem would be to generalize the expert policy $\pi_E$ outside state examples provided in this dataset. A first approach, recently used in \cite{melo2010learning} and legitimated for finite-horizon problems in \cite{syed2010reduction}, is to frame this problem as learning a multi-classifier where each action $a \in A$ will be a class. Indeed, trained on the training set $\{s_i,a_i=\pi_E(s_i)\}_{1\leq i \leq D}$, many classifiers are able to output a score function $q$ from $S\times A$ to $\mathbb{R}$. The final decision rule $\pi_C$ which is a function from $S$ to $\mathbb{R}$ and thus can be seen as a deterministic policy is constructed in order to satisfy:

@@

Now, one can notice the similarity between \eqref{equation:Qoptimal} and \eqref{equation:Coptimal}. The score-function $q$ can then be interpreted as an action-value function and it is easy and natural to construct a reward function $R_C$ for which the policy $\pi_C$ is optimal as it will be proven in the section \ref{section: Analysis}. Let define $R^C$ as:

@@

However, the model being unknown (contrary to $q$ and $\pi_C$) it is not possible to compute $R^C$ directly. However, it can be estimated from the available transitions (see
\eqref{equation:data}). Let us write:

@@

\label{r.def}
\hat{r}_i=q(s_i,a_i)-\gamma q(s_i',\pi_C(s_i')), \forall i\in \{1,\dots,D\}.

@@

This is an unbiased estimate of $R_C(s_i,a_i)$, indeed:

@@

R_C(s_i,a_i)=\sum_{s'\in S}p(s_i,a_i,s')[q(s_i,a_i)-\gamma q(s',\pi_C(s'))], \forall i\in \{1,\dots,D\}.

@@

Therefore, it is possible to build an estimate $\hat{R}_C$ of $R_C$, using a regressor trained on the dataset:

@@

\{(s_i,a_i,s'_i,\hat{r}_i)\}_{1\leq i \leq D}.

@@

Finally, we see $\hat{R}_C$ as an approximation of $R_C$ and $\hat{\pi}_C$ will be an optimal policy for the reward $\hat{R}_C$. In order to verify that the reward function $\hat{R}_C$ is a good candidate to resolve the IRL problem, it will be proven in the section \ref{section: Analysis} that the policies $\pi_E$ and $\hat{\pi}_C$ are quite similar (confer lemma \ref{lemma : results}).

@@

where $E_{s \sim \mu_\pi}$ means that $s$ is distributed over the distribution $\mu_\pi$. All the results provided in the next theorem use the expectation $E_{s \sim \mu_E}$ which is the canonical expectation
to consider when one wants to prove a result over the state space $S$ related to the policy of the expert $\pi_E$.\\

@@

One can notice that the error classification is defined thanks to the expectation $E_{s \sim \mu_E}$. In the classical framework of classification the data $\{s_i,a_i=\pi_E(s_i)\}_{1\leq i \leq D}$
are generated independently according to a distribution $\mu_{\text{Data}}$ over $S$ and hence the classical classification error must be:

@@

Let $\{S,A,P,\gamma\}$ be a finite MDP without reward and $\pi_E$ an expert policy. The only available data is depicted in \eqref{equation:data}.
The notations $q$, $\pi_C$, $\hat{\pi}_C$, $\hat{R}^C$ are introduced in the section \ref{section: Cascading}.

@@

\item $\mu_E^T(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C})\leq \frac{\epsilon_C\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))+\|\epsilon^C_R\|_{\infty}+\|\epsilon^E_R\|_{\infty}}{1-\gamma}$.

@@

In order to understand why this theorem is useful, let us make some important assumptions. Let us suppose that the classification and the regression steps are perfect in the sense that $\epsilon_R(s,a)=0,\forall (s,a)\in S\times A$ and $\epsilon_C=0$. Then we obviously have, thanks to the theorem \ref{theorem : results}, that $\pi_E=\pi_C$ is optimal for $\hat{R}_C=R_C$. Thus the method is able to provide a non-trivial reward function for which the policy $\pi_E$ is optimal. Moreover if the classification step and the regression step are not perfect, the theorem \ref{theorem : results} shows, that our approach is able to provide a non trivial-reward $\hat{R}_C$ for which the policy $\pi_E$ is near-optimal in the sense that:

@@

E_{s \sim \mu_E}[V^{\hat{\pi}_C}_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq B[\epsilon_C+\|\epsilon_R\|_{\text{max}}], B\in\mathbb{R}_+,

@@

which means that the value functions of $\pi_E$ and $\hat{\pi}_C$ are not so different in the most visited states by the policy $\pi_E$.

@@

\begin{proof}[Lemma \ref{lemma: calculs V}]
Here, we use \eqref{equation: stationarity}.
\begin{align}
\mu_\pi^TV^\pi_R&=\mu_\pi^T(R_\pi+\gamma P_\pi V^\pi_R)=\mu_\pi^TR_\pi+ \gamma\mu_\pi^TP_\pi V^\pi_R,
\\
&=\mu_\pi^TR_\pi+ \gamma\mu_\pi^TV^\pi_R=\frac{1}{1-\gamma}\mu_\pi^TR_\pi.
\end{align}
\end{proof}

%\begin{remark}
%Let use the notations of the lemma \ref{lemma: calculs V}. For each $s\in S$, we define the average reward $\rho^\pi_R(s)$ such that :
%\begin{equation}
%\rho^\pi_R(s)=\lim_{n\rightarrow+\infty}\frac{1}{n}E^\pi_s[\sum_{t=0}^{n-1}\gamma^tR(s_t,\pi(s_t))]
%\end{equation}
%If $P_\pi$ is reducible then $\rho^\pi_R(s)$ does not depend on $s$ and:
%\begin{equation}
%\rho_R^\pi(s)=\rho_R^\pi=\mu_\pi^TR_\pi,\forall s\in S.
%\end{equation}
%The proof of that result is essentially an application of the ergodic theorem (see \cite{puterman1994markov}).
%So, if $P_\pi$ is reducible:
%\begin{equation}
%\mu_\pi^TV^\pi_R=\frac{1}{1-\gamma}\rho_R^\pi.
%\end{equation}
%There is a clear link between the average reward and the discounted reward.
%\end{remark}

\begin{proof}[Theorem \ref{lemma : results}]
In order to prove the tree results of the lemma \ref{lemma : results}, it is interesting to introduce the function $R_E$ from $S\times A$ to $\mathbb{R}$ such that:

@@

$\pi_C$ is optimal for the reward $R_C$.

@@

&\|V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}\|_{+\infty}\leq\frac{\|\epsilon^E_R\|_{\infty}}{1-\gamma},

@@

V^{\hat{\pi}_C}(s)_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}(s)&=E^{\hat{\pi}_C}_s[\sum_{t=0}^{+\infty}\gamma^t(\hat{R}^C(s_t,\hat{\pi}_C(s_t))-R^C(s_t,\hat{\pi}_C(s_t)))], \forall s\in S,

@@

@@

V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}(s)&=E^{\pi_E}_s[\sum_{t=0}^{+\infty}\gamma^t(R^C(s_t,\pi_E(s_t))-\hat{R}^C(s_t,\pi_E(s_t)))], \forall s\in S,
&\leq\frac{\|\epsilon^E_R\|_{\infty}}{1-\gamma}, \forall s\in S.

@@

\mu_E^T(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}+V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})\leq\frac{\|\epsilon^E_R\|_{\infty}+\|\epsilon^C_R\|_{\infty}}{1-\gamma}.

@@

\mu_E^TV^{\hat{\pi}_C}_{\hat{R}^C}-\mu_E^TV^{\pi_E}_{\hat{R}^C}&\leq \frac{\epsilon_C\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))+\|\epsilon^C_R\|_{\infty}+\|\epsilon^E_R\|_{\infty}}{1-\gamma},
&\leq\frac{\epsilon_C\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))+2\|\epsilon_R\|_{\text{max}}}{1-\gamma}.

>>>>

\title{Learning a reward function from demonstrations: a cascaded supervised
learning approach}
\author{Edouard Klein$^{1,2}$\\
 1. ABC Team\\
 LORIA-CNRS, France.
\And Bilal Piot$^{2}$\\
 2. Supélec-Metz Campus\\
 MaLIS Research group, France\\
\And Matthieu Geist$^1$\\
\texttt{prenom.nom@supelec.fr}\\
\And Olivier Pietquin$^{2,3}$\\
3. UMI 2958 CNRS\\
GeorgiaTech, France

@@

Given a sequential decision making problem posed in the Markov Decision Process (MDP) formalism, it is often difficult to specify manually a reward function in order to realize a given task. However, observing an expert demonstrating a task is often possible and easier~\cite{ng2000algorithms}. That is why learning from demonstrations and more particularly Inverse Reinforcement Learning (IRL) has received more and more interest this last decade. The IRL framework, first introduced in~\cite{russell1998learning},\cite{ng2000algorithms}, assumes that an expert, demonstrating a task, is acting optimally in a MDP with respect to an unknown reward function to be discovered. Unlike direct methods of learning from demonstrations which learn a mapping from states to actions via supervised learning(SL)~\cite{atkeson1997robot},\cite{pomerleau1989alvinn}, IRL methods aim at recovering a reward function which is the most compact and transferable way to specify a task. Moreover most of the direct methods do not use the structure of the MDP (one notable exception is~\cite{melo2010learning}), hence mimic the policy of the expert and are not able to generalize correctly the expert policy outside state examples. In order to overcome this drawback, undirect methods known as apprenticeship learning~\cite{abbeel2004apprenticeship} were introduced. These methods (see~\cite{neu2009training} for a comprehensive overview) uses the IRL framework but most of them assume the dynamic of the environment known, or an environment model (via a simulator) must be available so as to test the effects of a policy. Moreover, most of existing undirect approaches require to solve the direct reinforcement learning problem (find an optimal policy knowing a given reward) several times (one notable exception is~\cite{boularias2011relative}). Finally, in many cases, the output of the algorithm is a policy but not a reward function (see Section \ref{section: related work} for an in-depth presentation). These constraints can lead to the development of algorithms unapplicable to real-world applications and make the problem of apprenticeship learning harder than reinforcement learning. That is why a reborn interest for direct methods which passively imitate the expert exists. More particularly the reduction of the apprenticeship learning to classification, first introduced in~\cite{zadrozny2003cost} and recently used in~\cite{melo2010learning}, is legitimated for finite horizons problems in~\cite{syed2010reduction},\cite{ross2010efficient}. Our approach tries to avoid the drawbacks of the direct and undirect methods by cascading two well known supervised learning(see Section \ref{section: Cascading}): first a step of classification and then a step of regression which introduces the MDP structure and  outputs a reward function (see Section \ref{section: Cascading}). Our approach doesn't require any of the following: complete trajectories from an expert, a generative model of the environment, the knowledge of the transition probabilities, the ability to compute a (near)-optimal policy for different rewards and hence resolve the forward reinforcement learning (RL) problem, the perfect knowledge of the expert's policy. In addition, only a small amount of expert demonstrations (not even in the form of trajectories but simple transitions) is required.\\
The remainder of the paper is organized as follows: Section \ref{section: background} gives us the different notations used in the paper and presents briefly the RL and IRL frameworks, Section \ref{section: Cascading} present our method which is legitimated in Section \ref{section: Cascading}, Section \ref{section: experiments} presents an instantiation of our method and tests it on different benchmarks and finally Section \ref{section: related work} compares our method to the state of the art.

@@

A (finite) MDP \cite{puterman1994markov} is a tuple $M=\{S,A,P,\gamma,R\}$ where $S=\{s_i\}_{1\leq i \leq N}$ is the finite state space with $N\in\mathbb{N}^*$ states, $A=\{a_k\}_{1\leq k \leq K}$ is the finite action space with $K\in\mathbb{N}^*$ actions, $P=\{P_{s,a}\}_{(s,a)\in S\times A}$ is the set of transition probabilities where $P_{s,a}$ is a distribution probability over $S$, $P$ represents the dynamics of the MDP and the notation $p(s,a,s')=P_{s,a}(s')$ is often used and quantifies the probability to reach $s'\in S$ knowing that the action $a \in A$ was taken in the state $s\in S$, $\gamma\in]0,1[$ is the discount factor and $R$ is a function from $S\times A$ to $\mathbb{R}$ called the reward function. A deterministic policy $\pi$ is a function from $S$ to $A$, $\Pi$ is the set of deterministic policies and the value function $V^\pi_R$ is a function from $S$ to $\mathbb{R}$ defined by:

@@

For a finite-MDP $M=\{S,A,P,\gamma,R\}$, the existence of an optimal policy for any function $R$ is shown in \cite{bertsekas2001dynamic}.

@@

We will also need the following matrices and column-vectors defined for any dynamic $P$, any policy $\pi$ and reward function $R$ :$P_a=[P_{s_i,a}(s_j)]_{1\leq i,j \leq N}$, $P_\pi=[P_{s_i,\pi(s_i)}(s_j)]_{1\leq i,j \leq N}$, $R_\pi=[R(s_i,\pi(s_i))]_{1\leq i,j \leq N}$. The transposition of a real column-vector $X=[X_i]_{1\leq i \leq I}$, $I\in\mathbb{N}^*$, is the real row-vector $X^T$ and $\|X\|_{\infty}=\max_{1\leq i \leq I}|X_i|$. The transposition of the real matrice $Y=[Y_{i,j}]_{1\leq i \leq I, 1\leq j \leq J}$, $(I,J)\in\mathbb{N}^2$, will be $Y^T$ and $\|Y\|_{\text{max}}=\max_{1\leq i \leq I,1\leq j \leq J}|Y_{i,j}|$.  Thanks to these vectorial notations, we can define the Bellman operator $T^\pi_R$, for a given deterministic policy $\pi$ and a given reward $R$, which is a function from $\mathbb{R}^N$ to $\mathbb{R}^N$ such that:

@@

representation), that the model ($P$ and $R$) may be unknown (the only information being provided through rewarded transitions sampled according
to some behavioral policy), that learning can occurs in an online or off-line setting, and so on. There exists books on the subject \cite{bertsekas2001dynamic},\cite{sutton1998reinforcement}.\\
In the classical IRL paradigm \cite{ng2000algorithms}, an MDP without reward $M\backslash R =\{S,A,P,\gamma\}$ and a policy $\pi_E$ called expert-policy are given and the problem is to find
a reward $R^*$ for which the policy $\pi_E$ is optimal. However this problem is clearly ill-posed in the sense that there is not uniqueness of the reward $R^*$ : many rewards functions are equivalent in the sense that they have the same optimal deterministic policies \cite{ng1999policyreward}, moreover the trivial zero-reward is a solution for any deterministic policy $\pi_E$ as it is shown in~\cite{ng2000algorithms}. In the literature, some solutions are proposed in order to respond to the ill-posed nature of the problem \cite{ng2000algorithms},\cite{ziebart2008maximum},\cite{boularias2011relative}.
In our experiments, see Section \ref{section: experiments}, we assume that the solely available information is provided by transitions sampled according to the dynamics of the environnement under $\pi_E$:

@@

\{(s_i,a_i=\pi_E(s_i)),s_i')\}_{1\leq i \leq D}, D\in\mathbb{N}^*,

@@

where $s_i'$ is sampled according to the distribution $P_{s_i,a_i}$.

@@

For a given deterministic policy $\pi$, a component $P_{s_i,\pi(s_i)}(s_j)$ of the matrix $P_\pi$ represents the probability to transit from $s_i$ to $s_j$ under the policy $\pi$. So $P_\pi$ can be seen as a transition matrix of a finite Markov-chain on the finite state space $S$. Thus, let us recall some basic but important results of finite-Markov chain theory.

@@

\begin{theorem}[\cite{baldi2000martingales}]

@@

The available data, provided by the expert, is being depicted by:
D_C=\{(s_{C,i},a_{C,i}=\pi_E(s_{C,i})),s'_{C,i})\}_{1\leq i \leq D}, D\in\mathbb{N}^*,
where $s'_{C,i}$ is sampled according to the distribution $P_{s_{C,i},a_{C,i}}$.
A first problem would be to generalize the expert policy $\pi_E$ outside state examples provided in this dataset. A first approach, recently used in \cite{melo2010learning} and legitimated for finite-horizon problems in \cite{syed2010reduction}, is to frame this problem as learning a multi-classifier where each action $a \in A$ will be a class. Indeed, trained on the training set $\{s_{C,i},a_{C,i}=\pi_E(s_{C,i})\}_{1\leq i \leq D}$, many classifiers are able to output a score function $q$ from $S\times A$ to $\mathbb{R}$. The final decision rule $\pi_C$ which is a function from $S$ to $\mathbb{R}$ and thus can be seen as a deterministic policy is constructed in order to satisfy:

@@

Now, one can notice the similarity between equation~\eqref{equation:Qoptimal} and equation~\eqref{equation:Coptimal}. The score-function $q$ can then be interpreted as an action-value function and it is easy and natural to construct a reward function $R^C$ for which the policy $\pi_C$ is optimal as it will be proven in the section~\ref{section: Analysis}. Let define $R^C$ as:

@@

However, the model being unknown (contrary to $q$ and $\pi_C$) it is not possible to compute $R^C$ directly. But, it can be estimated from transitions of a regression data set:

@@

D_R=\{(s_{R,i},a_{R,i},s'_{R,i})\}_{1\leq i \leq D'}, D'\in\mathbb{N}^*,

@@

where $s'_{R,i}$ is sampled under the probability $P_{s_{R,i},a_{R,i}}$.
Let us write:

@@

\label{ri.def}
\hat{r}_i=q(s_{R,i},a_{R,i})-\gamma q(s'_{R,i},\pi_C(s'_{R,i})), \forall i\in \{1,\dots,D'\}.

@@

This is an unbiased estimate of $R^C(s_{R,i},a_{R,i})$, indeed:

@@

R^C(s_{R,i},a_{R,i})=\sum_{s'\in S}p(s_{R,i},a_{R,i},s')[q(s_{R,i},a_{R,i})-\gamma q(s',\pi_C(s'))], \forall i\in \{1,\dots,D'\}.

@@

Therefore, it is possible to build an estimate $\hat{R}^C$, which is a function from $S\times A$ to $\mathbb{R}$, of $R^C$ using a regressor trained on the dataset:
\{(s_{R,i},a_{R,i},\hat{r}_i)\}_{1\leq i \leq D'}.
Finally, we see $\hat{R}^C$ as an approximation of $R^C$ and $\hat{\pi}_C$ is defined as an optimal policy for the reward $\hat{R}^C$. In order to verify that the reward function $\hat{R}^C$ is a good candidate to resolve the IRL problem, it will be proven in Section \ref{section: Analysis} that the policy $\pi_E$ is near-optimal for the reward $\hat{R}^C$ (confer theorem \ref{theorem : results}).
\subsection{A particular case}
An interesting case is  when $D_C=D_R$ which is the one chosen in Section \ref{section: experiments}.
In this case, the solely available data are the expert data depicted by the equation \ref{equation:data}.
The approximation function $\hat{R}^C$ obtained thanks to the regression is only accurate on the set $(s,\pi_E(s))_{\{s\in S\}}$ because of the
regression training set.

@@

where $E_{s \sim \mu_\pi}$ means that $s$ is distributed over the distribution $\mu_\pi$. All the results provided in the next theorem use the expectation $E_{s \sim \mu_E}$ which is the canonical expectation to consider when one wants to prove a result over the state space $S$ related to the policy of the expert $\pi_E$.\\

@@

One can notice that the classification error is defined thanks to the expectation $E_{s \sim \mu_E}$. In the classical framework of classification, the data $\{s_i,a_i=\pi_E(s_i)\}_{1\leq i \leq D}$ are generated independently according to a distribution $\mu_{\text{Data}}$ over $S$ and hence the classical classification error must be:

@@

Let $\{S,A,P,\gamma\}$ be a finite MDP without reward and $\pi_E$ an expert policy.
The notations $q$, $\pi_C$, $\hat{\pi}_C$, $\hat{R}^C$ are introduced in the Section \ref{section: Cascading}.

@@

\item $\mu_E^T(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C})\leq \frac{\epsilon_C\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))+\|\epsilon^C_R\|_{\infty}+\mu_E^T\epsilon^E_R}{1-\gamma}$.

@@

In order to understand why this theorem is useful, let us make some important assumptions. Let us suppose that the classification and the regression steps are perfect in the sense that $\epsilon_R(s,a)=0,\forall (s,a)\in S\times A$ and $\epsilon_C=0$. Then we obviously have, thanks to the theorem \ref{theorem : results}, that $\pi_E=\pi_C$ is optimal for $\hat{R}^C=R^C$. Thus the method is able to provide a non-trivial reward function for which the policy $\pi_E$ is optimal. Moreover if the classification step and the regression step are not perfect, the theorem \ref{theorem : results} shows, that our approach is able to provide a non trivial-reward $\hat{R}^C$ for which the policy $\pi_E$ is near-optimal in the sense that:

@@

E_{s \sim \mu_E}[V^{\hat{\pi}_C}_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq B[\epsilon_C+\|\epsilon^C_R\|_{\infty}+\mu_E^T\epsilon^E_R], B\in\mathbb{R}_+.

@@

It is important to be clear about this result. If the only available data are provided by the equation \eqref{equation:data}, it is possible to control
$\epsilon_C$ and $\mu_E^T\epsilon^E_R$ because these errors depend only on the expert policy $\pi_E$. However it is not possible to control the error $\|\epsilon^C_R\|_{\infty}$
because it depends on the policy $\hat{\pi}_C$ which can be different than the expert policy and hence do not appear in the available data \eqref{equation:data}. However
it will be possible to obtain a control on the term $\|\epsilon^C_R\|_{\infty}$ if the data used for the regression are:
D_R=\{(s_{R,i},a_{R,i},s'_{R,i})\}_{1\leq i \leq D'},
where $(s_{R,i},a_{R,i})$ are uniformly chosen on the set $S\times A$ or sampled from other policies than the expert. So, theoretically an easy way to be sure
to control the error $\|\epsilon^C_R\|_{\infty}$ is to be able to give a data set for the regression which is sampled from the expert policy and other policies (and more particularly $\hat{\pi}_C$). But we give examples, see Section \ref{section: experiments}, where the regression data set given by the equation \eqref{equation:data} is sufficient to obtain good results.
A possible argument to explain the fact that classification-regression still works when $D_C=D_R$, is that $\hat{\pi}_C$ must not be so different than $\pi_E$ but we did not manage to control
the term $\|\epsilon^C_R\|_{\infty}$ when $D_C=D_R$.

@@

\begin{proof}[Lemma \ref{lemma: calculs V}]
Here, we use \eqref{equation: stationarity}
\mu_\pi^TV^\pi_R&=\mu_\pi^T(R_\pi+\gamma P_\pi V^\pi_R)=\mu_\pi^TR_\pi+ \gamma\mu_\pi^TP_\pi V^\pi_R,
&=\mu_\pi^TR_\pi+ \gamma\mu_\pi^TV^\pi_R=\frac{1}{1-\gamma}\mu_\pi^TR_\pi.
\begin{proof}[Theorem \ref{theorem : results}]
In order to prove the three results of the theorem \ref{theorem : results}, let us introduce the function $R_E$ from $S\times A$ to $\mathbb{R}$ such that:

@@

$\pi_C$ is optimal for the reward $R^C$.

@@

&\mu_E^T(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})\leq\frac{\mu_E^T\epsilon^E_R}{1-\gamma},

@@

V^{\hat{\pi}_C}_{\hat{R}^C}(s)-V^{\hat{\pi}_C}_{R^C}(s)&=E^{\hat{\pi}_C}_s[\sum_{t=0}^{+\infty}\gamma^t(\hat{R}^C(s_t,\hat{\pi}_C(s_t))-R^C(s_t,\hat{\pi}_C(s_t)))], \forall s\in S,

@@

And, thanks to the lemma \ref{lemma: calculs V}:
\mu_E^T(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})&=\mu_E^T(V^{\pi_E}_{\epsilon_R}),

@@

&\leq\frac{\mu_E^T\epsilon^E_R}{1-\gamma}.

@@

\mu_E^T(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}+V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})\leq\frac{\mu_E^T\epsilon^E_R+\|\epsilon^C_R\|_{\infty}}{1-\gamma}.

@@

\mu_E^TV^{\hat{\pi}_C}_{\hat{R}^C}-\mu_E^TV^{\pi_E}_{\hat{R}^C}\leq \frac{\epsilon_C\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))+\|\epsilon^C_R\|_{\infty}+\mu_E^T\epsilon^E_R}{1-\gamma}.
